---
title: "MAF_tools_rm_BT972"
output: html_notebook
---


```{r}
if (!require("BiocManager", quietly = TRUE))
     install.packages("BiocManager")
if (!require("maftools", quietly = TRUE))
     install.packages("maftools")
```

```{r}
library(maftools)


my_oncoplot <- function (maf_file, clin_data, samples.vector, title="Top 30 genes altered", file.out="plot_name"){
 
 #Fields found in MAF file. They depend on Open Cravat. 
 vc_nonSyn_vector <- c("", "synonymous_variant", "splice_site_variant", "2kb_downstream_variant", "frameshift_truncation", "frameshift_elongation", "inframe_insertion", "miRNA", "misc_RNA", "polymorphic_pseudogene", "3_prime_UTR_variant", "stop_gained", "inframe_deletion", "snRNA", "2kb_upstream_variant", "missense_variant", "NSD_transcript", "IG_V_gene", "rRNA", "5_prime_UTR_variant", "start_lost", "stop_lost", "stop_retained_variant", "processed_transcript", "sRNA", "NMD_transcript_variant", "TR_C_gene", "lnc_RNA", "snoRNA", "intron_variant")
 # This reads the file in MAF-tools.
 laml <- read.maf(maf = maf_file,vc_nonSyn = vc_nonSyn_vector, clinicalData = clin_data)
 plotmafSummary(maf = laml, rmOutlier = TRUE,  dashboard = TRUE, titvRaw = FALSE)
 
 annotation_colors <- list(Treatment_Group = c(sensitive = "#D53E4F", resistant = "#F46D43"))
 
 maf_filtered <- laml
 #maf_filtered@data <- maf_filtered@data[maf_filtered@data$Hugo_Symbol != "Unknown", ]
 
 maf_filtered <- subsetMaf(maf = maf_filtered, tsb = samples.vector)
 
 #Saves to png
 png(paste("./MAFTools/",file.out,".png"), width = 3000, height = 2000, res = 300)
 
 # Generate the oncoplot.
 oncoplot(
   maf = maf_filtered,
   top = 20,
   titleText = title,
   sortByAnnotation = TRUE,
   annotationColor = annotation_colors,
   showTumorSampleBarcodes = TRUE,
   drawRowBar = FALSE,
   barcodeSrt = 1,
   gene_mar = 8
   )
 dev.off()
 
 }

```


Mutations status if appear in 2 or more sensitive samples. Genes of interest for sensitive mutations could not be plotted.

```{r, width=4, height=10}
#Oncoplot of strict sensitive unique samples. 
clin_dataS <- "./MAFTools/rm_BT972_COMPLETE_MAF_tools_metadata.csv"
maf_fileS <- "./MAFTools/sensitive__strict_damaging.maf"

my_oncoplot(maf_file = maf_fileS, 
            clin_data =clin_dataS, 
            samples.vector = c("S_BT241","S_BT935","S_MBT168"),
            title= "Top 20 genes altered in sensitive cell lines",
            file.out="sensitive_strict_oncoplot")


#Oncoplot of top mutations of strict sensitive gene of interest. 
maf_file.interest.S <- "./MAFTools/sensitive_strict_genes_interest.maf"

my_oncoplot(maf_file = maf_file.interest.S, 
            clin_data =clin_dataS, 
            samples.vector = c("S_BT241","S_BT935","S_MBT168"),
            title= "Top 20 genes of interest altered in sensitive cell lines")
```


Mutations status if appear in 2 or more resistant samples. Genes of interest for resistant mutations are not investigated, as filtering yielded none.
```{r, width=4, height=10}
#Oncoplot of strict resistant unique samples.
clin_dataR <- "./MAFTools/rm_BT972_COMPLETE_MAF_tools_metadata.csv"
maf_fileR <- "./MAFTools/resistant__strict_damaging.maf"

my_oncoplot(maf_file = maf_fileR, 
            clin_data =clin_dataR, 
            samples.vector = c("R_BT594","R_MBT357","R_MBT373"), 
            title= "Top 20 genes altered in resistant cell lines",
            file.out="resistant_strict_oncoplot")


#Oncoplot of top mutations of strict resistant gene of interest. 
maf_file.interestR <- "./MAFTools/"

my_oncoplot(maf_file = maf_file.interestR, 
            clin_data =clin_dataR, 
            samples.vector = c("R_BT594","R_MBT357","R_MBT373"),  
            title= "Top 30 genes of interest altered in resistant cell lines")


```



### T-test of total and damaging mutational burden 

H0: there is no difference in mutation amount between the two groups
H1: there is a difference in mutation amount between the two groups

```{r}
library(ggplot2)
library(tidyr)
library(dplyr)

# Function for TMB statistical analysis
statistical_analysis <- function(resistant, sensitive, plot_title=""){
 count_TR <- log(resistant$count)
 resistant$count <- count_TR
 resistant_summaryT <- c(median(resistant$count),max(resistant$count),min(resistant$count)) 
 
 
 count_TS <- log(sensitive$count)
 sensitive$count <- count_TS
 sensitive_summaryT <- c(median(sensitive$count),max(sensitive$count),min(sensitive$count)) 
 
 
 # Variance not met, using Mann Whitney
 result_T <- wilcox.test(count_TS, count_TR)
 print(result_T)
 
 summary_df_T <- data.frame(sensitive = sensitive_summaryT, resistant=resistant_summaryT , row.names = c("median", "max","min"))
 
 data_long <- pivot_longer(summary_df_T, cols = c("sensitive", "resistant"), names_to = "group", values_to = "value")
 
 # Create the whisker plot
 fig <- ggplot(data_long, aes(x = group, y = value)) +
   geom_boxplot(aes(group = group)) +
   geom_text(aes(label = round(value, 2)), 
             position = position_dodge(width = 0.75), 
             vjust = -0.5, 
             size = 3) +
   labs(title = paste("Logged counts of", plot_title, "mutations depending on cell line"),
        x = "Cell line",
        y = "Logged mutation count") +
   theme_minimal()
 return (fig)
}
```


```{r}

### Loading in the datasets.
total_mutations <- read.csv("./rm_BT972_COMPLETE_total.csv")

damaging_mutations <- read.csv("./mutations_ont_summary.csv")



### Total mutations analysis. Counts are logged.

resistant_T <- subset(total_mutations, samples == c("R_BT594","R_MBT357", "R_MBT373"))
sensitive_T <- subset(total_mutations, samples == c("S_BT241","S_BT935", "S_MBT168"))

var.test(log(sensitive_T$count),log(resistant_T$count)) 

figure_T <- statistical_analysis(resistant_T,sensitive_T, plot_title = "total")

ggsave("./Total_mutations_whisk.png", plot = figure_T, width = 6, height = 4)
```


```{r}
### Damaging mutation analysis. Counts are logged.
damaging_mutations <- damaging_mutations[c("Sample_Name","total.mutations")]
colnames(damaging_mutations) <- c("samples", "count")

sensitive_D <- subset(damaging_mutations, samples == c("S_BT241","S_BT935", "S_MBT168"))
resistant_D <- subset(damaging_mutations, samples == c("R_BT594","R_MBT357", "R_MBT373"))

var.test(log(sensitive_D$count),log(resistant_D$count)) # Variance not. Using Wilcox test. 


figure_D <- statistical_analysis(resistant = resistant_D, sensitive = sensitive_D, plot_title = "damaging")

ggsave("./Damaging_mutations_whisk.png", plot = figure_D, width = 6, height = 4)
```



The next section focuses on comparing the statistical significance in the difference between the unique damaging mutations in each group. 
```{r}
sens_counts.df <- as.data.frame(read.csv("./MAFTools/Sensitive_summary_count.csv"))
res_counts.df <- as.data.frame(read.csv("./MAFTools/Resistant_summary_count.csv"))

sens_counts.vec <- as.vector(sens_counts.df[,"Counts"])
res_counts.vec <- as.vector(res_counts.df[,"Counts"])

var.test(sens_counts.vec,res_counts.vec) # Homeodascity of variance met.

t.test(sens_counts.vec,res_counts.vec)


boxplot(sens_counts.vec, res_counts.vec, 
        names = c("Sensitive cell line", "Resistant cell line"), 
        main = "Unique mutation counts", ylab = "Mutation Counts")

```



